---
title: "Data extraction from xml inR"
author: "Jai Iyyappa Vignesh Manivannan"
output:
  html_document:
    df_print: paged
  pdf_document: default
---



```{r}
# Installing the RMySQL Library
library(RMySQL)


#  MySQL Settings
db_user <- 'dbuser'
db_password <- 'dbpass'
db_name <- 'dbname'
db_host <- 'hostname'
db_port <- 3306

# Establishing a connection to MySQL
mydb <-  dbConnect(MySQL(), user = db_user, password = db_password,
                  dbname = db_name, host = db_host, port = db_port)


```


```{r loadLibraries}
library(XML)
```

### The XML file path is given.
```{r loadPath}
path <- "C:/Users/Jai Vignesh M/Documents/DBMS Practicum 2/pubmed_sample.xml"
```

### Parse and load XML into Document Object Model (DOM)

```{r}
# Reading the XML file and parse into DOM
xmlDOM <- xmlParse(file = path)
```

### Determining the root element

```{r}
r <- xmlRoot(xmlDOM)
# number of children of root (number of PubmedArticle)
totalRow <- xmlSize(r)
```
### Initializing all the data frames

#### Article data frame
```{r}
Article.df <- data.frame (articleId = vector (mode = "integer", 
                                         length = totalRow),
                         articleTitle = vector (mode = "character", 
                                         length = totalRow),
                         articleLanguage = vector (mode = "character", 
                                         length = totalRow),
                         articleDate = vector (mode = "character", 
                                         length = totalRow),
                         articlePubModel = vector (mode = "character", 
                                         length = totalRow),
                         issn = vector (mode = "character", 
                                         length = totalRow),
                          stringsAsFactors = F)
```

#### Journal data frame

```{r}
Journal.df <- data.frame (issn = character(),
                          issnType = character(),
                          title = character(),
                          isoAbbreviation = character(),
                          journalIssueVolume = integer(),
                          journalIssueNum = integer(),
                          journalPublishDate = character(),
                          journalPublishYear = integer(),
                          journalPublishMonth = character(),
                          stringsAsFactors = F)
```


#### Author data frame

```{r}
Author.df <- data.frame (authorId = integer(),
                         authorLastName = character(),
                          authorForeName = character(),
                          authorInitial = character(),
                         authorAffiliation = character(),
                          stringsAsFactors = F)
```


#### ArticleAuthorMap data frame


```{r}
ArticleAuthorMap.df <- data.frame (authorArticleMapId = integer(),
                                   articleId = integer(),
                                   authorId = integer(),
                          stringsAsFactors = F)
```


#### MedlineJournalInfo data frame

```{r}
MedlineJournalInfo.df <- data.frame (infoId = integer(),
                                     articleId = integer(),
                                     country = character(),
                                     medlineTA = character(),
                                     nlmUniqueID = integer(),
                                     issnLinking = character(),
                            stringsAsFactors = F)
```


#### PubMedHistory data frame


```{r}
PubMedHistory.df <- data.frame (historyId = integer(),
                                articleId = integer(),
                                pubMedDateType = character(),
                                pubMedDate = character(),
                                stringsAsFactors = F)
```


### Parsing node data into data frame

#### parse Journal nodes

```{r parseJournal}
parseJournal <- function (aJournalNode)
{
  # parse the address into its components
  issn <- xmlValue(aJournalNode[[1]])
  issnType <- xmlAttrs(aJournalNode[[1]])[1]
  title <- xmlValue(aJournalNode[[3]])
  isoAbbreviation <- xmlValue(aJournalNode[[4]])
  journalPublishDate <- ""
  journalIssueVolume <- xmlValue(aJournalNode[[2]][[1]])
  journalIssueNum <- xmlValue(aJournalNode[[2]][[2]])
  if(is.integer(aJournalNode[[2]][[3]][[1]]) && xmlName(aJournalNode[[2]][[3]][[1]]) == "Year") {
    journalPublishYear <- xmlValue(aJournalNode[[2]][[3]][[1]])
  }
  else{
    journalPublishYear <- substr(xmlValue(aJournalNode[[2]][[3]][[1]]),1,4)
  }
  
  if(!is.null(aJournalNode[[2]][[3]][[2]]) && xmlName(aJournalNode[[2]][[3]][[2]]) == "Month") {
    journalPublishMonth <- xmlValue(aJournalNode[[2]][[3]][[2]])
  }
  else{
    journalPublishMonth <- substr(xmlValue(aJournalNode[[2]][[3]][[1]]),6,8)
  }
  
  
  newJournal.df <- data.frame(issn, issnType, title, isoAbbreviation, 
                           journalIssueVolume, journalIssueNum,
                           journalPublishDate, journalPublishYear,
                           journalPublishMonth,  
                           stringsAsFactors = F)
  
  return(newJournal.df)
}
```

#### parse Author nodes

```{r parseAuthor}
parseAuthor <- function (anAuthorNode)
{
  # parse the address into its components
  authorLastName <- xmlValue(anAuthorNode[[1]])
  authorForeName <- xmlValue(anAuthorNode[[2]])
  authorInitial <- xmlValue(anAuthorNode[[3]])
  authorAffiliation <- xpathSApply(anAuthorNode, "./Affiliation", xmlValue)
    if (length(authorAffiliation) == 0)
      authorAffiliation <- ""
 
  newAuthor.df <- data.frame(authorLastName, authorForeName, authorInitial,authorAffiliation,
                           stringsAsFactors = F)
  
  return(newAuthor.df)
}
```

#### parse parseMedlineJournal nodes

```{r parseMedlineJournal}
parseMedlineJournalInfo <- function (aninfoNode)
{
  # parse the address into its components
  country <- xmlValue(aninfoNode[[1]])
  medlineTA <- xmlValue(aninfoNode[[2]])
  nlmUniqueID <- xmlValue(aninfoNode[[3]])
  issnLinking <- xmlValue(aninfoNode[[4]])
 
  newJournalInfo.df <- data.frame(country, medlineTA, nlmUniqueID, issnLinking,
                           stringsAsFactors = F)
  
  return(newJournalInfo.df)
}
```

### Functions for checking unique data

#### Functions for checking unique data based on entire content

```{r rowExists}
rowExists <- function (aRow, aDF)
{
  
  # check if that address is already in the data frame
  n <- nrow(aDF)
  c <- ncol(aDF)
  
  if (n == 0)
  {
    # data frame is empty, so can't exist
    return(0)
  }
  
  for (a in 1:n)
  {
    # check if all columns match for a row; ignore the aID column
    if (all(aDF[a,] == aRow[1,]))
    {
      # found a match; return it's ID
      return(a)
    }
  }
  
  # none matched
  return(0)
}
```

#### Functions for checking unique data based on primary key

```{r rowExistsPrimary}
rowPrimaryExists <- function (aRow, aDF)
{
  # check if that address is already in the data frame
  n <- nrow(aDF)
  c <- ncol(aDF)
  
  if (n == 0)
  {
    # data frame is empty, so can't exist
    return(0)
  }
  
  for (a in 1:n)
  {
    # check if all columns match for a row; ignore the aID column
    if (aDF[a,1] == aRow[1,1])
    { 
      # found a match; return it's ID
      return(a)
    }
  }
  
  # none matched
  return(0)
}
```

### Prepare author list and insert unique authors into data frame

```{r prepareAuthorList}
xpathExAuthor <- "//PubmedArticleSet/PubmedArticle/MedlineCitation/Article/AuthorList"
authorRowList <- xpathApply(xmlDOM,xpathExAuthor)

numAuthorList <- xmlSize(authorRowList)

index <- 1;

for (i in 1:numAuthorList)
{
  AuthorList <- authorRowList[[i]]
  
  xpathExAuthorInd <- "./Author"
  
  authorRows <- xpathApply(AuthorList,xpathExAuthorInd)

  numAuthors <- xmlSize(authorRows)
 

  for (j in 1:numAuthors)
{
  authorIndiv <- parseAuthor(authorRows[[j]])
   
   pk.authorUnique <- rowExists(authorIndiv, Author.df[,2:ncol(Author.df)])
   
   if(pk.authorUnique == 0){
     # does not exist, so add
    pk.authorUnique <- nrow(Author.df) + 1
    Author.df[pk.authorUnique,2:ncol(Author.df)] <- authorIndiv[1,]
    Author.df[pk.authorUnique,1] <- pk.authorUnique
   }
  
  }
  
}
```


### Parsing and Preparing all needed data into data frame

```{r}

# Reading the XML file and parse into DOM

# get the root node of the DOM tree
root <- xmlRoot(xmlDOM)

numArticlesRoots <- xmlSize(root)

# counters for index increase.
indexJournal <- 1
indexArticleAuthorMap <- 1
indexHistory <- 1

for (i in 1:numArticlesRoots)
{
  MedlineCitationData <- root[[i]][[1]]
  PubmedDataData <- root[[i]][[2]]
  
  
  #article id
  Article.df$articleId[i] <- i
  
   #article date
  articleDateYear <- xpathApply(MedlineCitationData,"./DateCreated/Year", xmlValue)
  articleDateMonth <- xpathApply(MedlineCitationData,"./DateCreated/Month", xmlValue)
  articleDateDate <- xpathApply(MedlineCitationData,"./DateCreated/Day", xmlValue)
  
  articleDate <- paste0(articleDateYear,"/",articleDateMonth,"/",articleDateDate)
  
  articleDate <- as.character.Date(articleDate)
  Article.df$articleDate[i] <- articleDate
  
  #journal publish date
  JournalDateYear <- xpathApply(PubmedDataData,"./History/PubMedPubDate[@PubStatus = 'pubmed']/Year", xmlValue)
  JournalDateMonth <- xpathApply(PubmedDataData,"./History/PubMedPubDate[@PubStatus = 'pubmed']/Month", xmlValue)
  JournalDateDate <- xpathApply(PubmedDataData,"./History/PubMedPubDate[@PubStatus = 'pubmed']/Day", xmlValue)
  
  journalDate <- paste0(JournalDateYear,"/",JournalDateMonth,"/",JournalDateDate)
  
  journalDate <- as.character.Date(journalDate)
  
 
  #journal publish date - checks implemented as instructed in the xml documentation
  if(as.numeric(JournalDateYear) - as.numeric(articleDateYear) > 0 && as.numeric(JournalDateMonth) >= as.numeric(articleDateMonth)) {
    journalDate <- as.character.Date(journalDate)
  }
  else{
    journalDate <- articleDate
  }
  
   
  #Article root
  articleList <- MedlineCitationData[[5]]
  
  if (xmlName(articleList) != "Article") {
    articleList <- MedlineCitationData[[4]]
  }
  
  #article Title
  articleTitle <- articleList[[2]]
  
  Article.df$articleTitle[i] <- xmlValue(articleTitle)
  
  #article Lang
  articleLanguage <- articleList[[7]]
  
  if(xmlName(articleLanguage) != "Language"){
     articleLanguage <- articleList[[8]]
  }
  
  Article.df$articleLanguage[i] <- xmlValue(articleLanguage)
  
  #article PubModel
  articleAttr <- xmlAttrs(articleList)
  
  Article.df$articlePubModel[i] <- articleAttr[1]
  
  # JOURNAL STEPS STARTS
  #Journal
  currentJournal <- articleList[[1]]
  
  #article ISSN
  articleIssn <- currentJournal[[1]]
  Article.df$issn[i] <- xmlValue(articleIssn)
  
  # Journal details START
  
  journalIndiv <- parseJournal(currentJournal)
  
  
  pk.journalUnique <- rowPrimaryExists(journalIndiv, Journal.df[,1:ncol(Journal.df)])
   
   
  if(pk.journalUnique == 0){
     Journal.df[indexJournal,1:ncol(Journal.df)] <- journalIndiv[1,]
     Journal.df[indexJournal,7] <- journalDate
     indexJournal = indexJournal+1;
  }
   
  

  # JOURNAL STEPS ENDS
  
  # Add mapping table between author and article
  
  authorList <- articleList[[7]]
  
  if(xmlName(authorList) != "AuthorList"){
     authorList <- articleList[[6]]
  }
  
  articleAuthorCount <- xmlSize(authorList)
  
  
  for(j in 1:articleAuthorCount){
    
    currentRowAuthor <- parseAuthor(authorList[[j]])
    
    pk.authorUnique <- rowExists(currentRowAuthor,Author.df[,2:ncol(Author.df)])
    
    newArticleAuthorMap.df <- data.frame(indexArticleAuthorMap, i, pk.authorUnique,
                           stringsAsFactors = F)
    
    ArticleAuthorMap.df[indexArticleAuthorMap,1:ncol(ArticleAuthorMap.df)] <- newArticleAuthorMap.df[1,]
    
    indexArticleAuthorMap = indexArticleAuthorMap + 1
    
  }
  
  # PUBLISHING STATUS STARTS for History
  
  historyList <- PubmedDataData[[1]]
  
  currentHistoryListCount <- xmlSize(historyList)
  
  
  for(x in 1:currentHistoryListCount){
    
    pubStatus <- xmlAttrs(historyList[[x]])[[1]]
    
     year <- xmlValue(historyList[[x]][[1]])
     month <- xmlValue(historyList[[x]][[2]])
     day <- xmlValue(historyList[[x]][[3]])
     
     date <- paste0(year,"/",month,"/",day)
     

     newPubStatus.df <- data.frame(indexHistory, i, pubStatus, date,
                           stringsAsFactors = F)
     
     PubMedHistory.df[indexHistory,1:ncol(PubMedHistory.df)] <- newPubStatus.df[1,]
     indexHistory = indexHistory+1;
     
  
    
  }
  
  
  # PUBLISHING STATUS STARTS for History ENDS
  
   #journalInfoList root
  journalInfoList <- MedlineCitationData[[6]]
  
  if (xmlName(journalInfoList) != "MedlineJournalInfo") {
    journalInfoList <- MedlineCitationData[[5]]
  }
  
  
  # medlineJournalInfo
  medJournalInfo <- parseMedlineJournalInfo(journalInfoList)
  
    MedlineJournalInfo.df[i,3:ncol(MedlineJournalInfo.df)] <- medJournalInfo[1,]
    MedlineJournalInfo.df[i,1] <- i
    MedlineJournalInfo.df[i,2] <- i
 
}


```

### TABLE CREATION

### Table creation based on the ER diagram



```{sql connection=mydb}
DROP TABLE IF EXISTS authorArticleMap
```


```{sql connection=mydb}
DROP TABLE IF EXISTS PubMedHistory
```


```{sql connection=mydb}
DROP TABLE IF EXISTS Article
```


#### Journal Table


```{sql connection=mydb}
DROP TABLE IF EXISTS Journal
```


```{sql connection=mydb}
CREATE TABLE Journal(
issn VARCHAR(25) NOT NULL,
issnType VARCHAR(255),
title VARCHAR(255),
isoAbbreviation VARCHAR(255),
journalIssueVolume INT,
journalIssueNum INT,
journalPublishDate DATE,
journalPublishYear INT,
journalPublishMonth VARCHAR(255),
PRIMARY KEY (issn)
)
```

#### Inserting the Journal data into table
```{r insertJournalDataIntoTable}
# dbWriteTable is used with append to insert the data from dataframe into table.
dbWriteTable(mydb, 'Journal', Journal.df, row.names=F, append=T, overwrite=F)
```
#### Display Journal table

```{sql connection=mydb}
SELECT * FROM Journal
       LIMIT 5;
```

#### Author Table

```{sql connection=mydb}
DROP TABLE IF EXISTS Author
```


```{sql connection=mydb}
CREATE TABLE Author(
authorId INT NOT NULL,
authorLastName VARCHAR(50),
authorForeName VARCHAR(50),
authorInitial VARCHAR(5),
authorAffiliation VARCHAR(250),
PRIMARY KEY (authorId)
)
```

#### Inserting the Author data into table
```{r insertAuthoDataIntoTable}
# dbWriteTable is used with append to insert the data from dataframe into table.
dbWriteTable(mydb, 'Author', Author.df, row.names=F, append=T, overwrite=F)
```


#### Display Author table

```{sql connection=mydb}
SELECT * FROM Author
       LIMIT 5;
```


#### Article Table

```{sql connection=mydb}
CREATE TABLE Article(
articleId INT NOT NULL,
articleTitle VARCHAR(255),
articleLanguage VARCHAR(50),
articleDate DATE,
articlePubModel VARCHAR(50),
issn VARCHAR(25) NOT NULL,
PRIMARY KEY (articleId),
FOREIGN KEY (issn) REFERENCES Journal(issn)
)
```

#### Inserting the Article data into table
```{r insertArticleDataIntoTable}
# dbWriteTable is used with append to insert the data from dataframe into table.
dbWriteTable(mydb, 'Article', Article.df, row.names=F, append=T, overwrite=F)
```


#### Display Article table

```{sql connection=mydb}
SELECT * FROM Article
       LIMIT 5;
```

#### authorArticleMap Table


```{sql connection=mydb}
CREATE TABLE authorArticleMap(
authorArticleMapId INT NOT NULL,
articleId INT NOT NULL,
authorId INT NOT NULL,
PRIMARY KEY (authorArticleMapId),
FOREIGN KEY (articleId) REFERENCES Article(articleId),
FOREIGN KEY (authorId) REFERENCES Author(authorId)
)
```

#### Inserting the ArticleAuthorMap data into table
```{r insertArticleAuthorMapDataIntoTable}
# dbWriteTable is used with append to insert the data from dataframe into table.
dbWriteTable(mydb, 'authorArticleMap', ArticleAuthorMap.df, row.names=F, append=T, overwrite=F)
```


#### Display authorArticleMap table

```{sql connection=mydb}
SELECT * FROM authorArticleMap
       LIMIT 5;
```

#### PubMedHistory Table

```{sql connection=mydb}
CREATE TABLE PubMedHistory(
historyId INT NOT NULL,
articleId INT NOT NULL,
pubMedDateType VARCHAR(255),
pubMedDate DATE,
PRIMARY KEY (historyId),
FOREIGN KEY (articleId) REFERENCES Article(articleId)
)
```

#### Inserting the PubMedHistory data into table
```{r insertPubMedHistoryDataIntoTable}
# dbWriteTable is used with append to insert the data from dataframe into table.
dbWriteTable(mydb, 'PubMedHistory', PubMedHistory.df, row.names=F, append=T, overwrite=F)
```


#### Display PubMedHistory table

```{sql connection=mydb}
SELECT * FROM PubMedHistory
       LIMIT 5;
```

#### MedlineJournalInfo Table


```{sql connection=mydb}
DROP TABLE IF EXISTS MedlineJournalInfo
```

                                     
```{sql connection=mydb}
CREATE TABLE MedlineJournalInfo(
infoId INT NOT NULL,
articleId INT NOT NULL,
country VARCHAR(255),
medlineTA VARCHAR(255),
nlmUniqueID INT,
issnLinking VARCHAR(255),
PRIMARY KEY (infoId)
)
```

#### Inserting the MedlineJournalInfo data into table
```{r insertMedlineJournalInfoDataIntoTable}
# dbWriteTable is used with append to insert the data from dataframe into table.
dbWriteTable(mydb, 'MedlineJournalInfo', MedlineJournalInfo.df, row.names=F, append=T, overwrite=F)
```


#### Display MedlineJournalInfo table

```{sql connection=mydb}
SELECT * FROM MedlineJournalInfo LIMIT 5;
```


### Creating the star/ snowflake - 


#### Dropping fact tables

#### ArticleFact DROP

```{sql connection=mydb}


DROP TABLE IF EXISTS starschema.ArticleFact;

```
#### SUMMARY TABLE DROP


```{sql connection=mydb}
DROP TABLE IF EXISTS starschema.SummaryFact;
```



### JournalDimension

```{sql connection=mydb}

DROP TABLE IF EXISTS starschema.JournalDimension

```

```{r}
# construct a SQL query
sqlCmdData1 = "CREATE TABLE starschema.JournalDimension 
   AS SELECT issn AS JournalDimensionKey,issnType,title,isoAbbreviation, journalIssueVolume, journalIssueNum, journalPublishDate FROM Practicum2DB.Journal"

tempData1 = dbGetQuery(mydb, sqlCmdData1)

sqlCmdaltData1 = "ALTER TABLE starschema.JournalDimension
   ADD CONSTRAINT JournalDimensionKey PRIMARY KEY (JournalDimensionKey);"

tempaltData1 = dbGetQuery(mydb, sqlCmdaltData1)
```

```{sql connection=mydb}
SELECT * FROM starschema.JournalDimension
       LIMIT 5;
```


### InformationArticleDimension

```{sql connection=mydb}

DROP TABLE IF EXISTS starschema.InformationArticleDimension;

```

```{r}
# construct a SQL query
sqlCmdData2 = "CREATE TABLE starschema.InformationArticleDimension 
   AS SELECT pa.articleId AS InformationArticleDimensionKey,pa.articleTitle,
   pa.articleLanguage,pa.articleDate, pa.articlePubModel, 
   pinfo.country,pinfo.medlineTA,pinfo.nlmUniqueID,pinfo.issnLinking
   FROM Practicum2DB.Article pa
   JOIN Practicum2DB.MedlineJournalInfo pinfo USING (articleId)"

tempData2 = dbGetQuery(mydb, sqlCmdData2)

sqlCmdaltData2 = "ALTER TABLE starschema.InformationArticleDimension
   ADD CONSTRAINT InformationArticleDimensionKey PRIMARY KEY (InformationArticleDimensionKey);"

tempaltData2 = dbGetQuery(mydb, sqlCmdaltData2)


```

```{sql connection=mydb}
SELECT * FROM starschema.InformationArticleDimension
       LIMIT 5;
```


### AuthorDimension

```{sql connection=mydb}

DROP TABLE IF EXISTS starschema.AuthorDimension;

```

```{r}
# construct a SQL query
sqlCmdData3 = "CREATE TABLE starschema.AuthorDimension 
   AS SELECT authorId AS authorDimensionKey,authorLastName,authorForeName,authorInitial FROM Practicum2DB.Author"

tempData3 = dbGetQuery(mydb, sqlCmdData3)

sqlCmdaltData3 = "ALTER TABLE starschema.AuthorDimension
   ADD CONSTRAINT authorDimensionKey PRIMARY KEY (authorDimensionKey);"

tempaltData3 = dbGetQuery(mydb, sqlCmdaltData3)
```

```{sql connection=mydb}
SELECT * FROM starschema.AuthorDimension
       LIMIT 5;
```


### AuthorMapDimension

```{sql connection=mydb}

DROP TABLE IF EXISTS starschema.AuthorMapDimension;

```

```{r}
# construct a SQL query
sqlCmdData4 = "CREATE TABLE starschema.AuthorMapDimension 
   AS SELECT authorArticleMapId AS mapAuthorDimensionKey,articleId AS InformationArticleDimensionKey,authorId AS authorDimensionKey FROM Practicum2DB.authorArticleMap"

tempData4 = dbGetQuery(mydb, sqlCmdData4)

sqlCmdaltData4 = "ALTER TABLE starschema.AuthorMapDimension
   ADD CONSTRAINT mapAuthorDimensionKey PRIMARY KEY (mapAuthorDimensionKey);"

tempaltData4 = dbGetQuery(mydb, sqlCmdaltData4)
```

```{sql connection=mydb}
SELECT * FROM starschema.AuthorMapDimension
       LIMIT 5;
```


### HistoryDimension

```{sql connection=mydb}

DROP TABLE IF EXISTS starschema.HistoryDimension;

```


```{r}
# construct a SQL query
sqlCmdData6 = "CREATE TABLE starschema.HistoryDimension 
   AS SELECT historyId AS historyDimKey,pubMedDateType,pubMedDate FROM Practicum2DB.PubMedHistory"

tempData6 = dbGetQuery(mydb, sqlCmdData6)


sqlCmdaltData6 = "ALTER TABLE starschema.HistoryDimension
   ADD CONSTRAINT historyDimKey PRIMARY KEY (historyDimKey);"

tempaltData6 = dbGetQuery(mydb, sqlCmdaltData6)
```

```{sql connection=mydb}
SELECT * FROM starschema.HistoryDimension
       LIMIT 5;
```


### PublishDateDimension

```{sql connection=mydb}

DROP TABLE IF EXISTS starschema.PublishDateDimension;

```


```{r}
# construct a SQL query for create
sqlCmdData7 = "CREATE TABLE starschema.PublishDateDimension (
  pubDatedim_key int NOT NULL AUTO_INCREMENT PRIMARY KEY,
  journalPublishDate DATE,
  day_of_week INT,
year_num INT,
quarter_of_year INT
)"

tempData7 = dbGetQuery(mydb, sqlCmdData7)

# construct a SQL query for insert
sqlCmdData8 = "INSERT INTO starschema.PublishDateDimension (journalPublishDate,day_of_week,year_num,quarter_of_year)
SELECT DISTINCT journalPublishDate,dayofweek(journalPublishDate) AS day_of_week,year(journalPublishDate) AS year, quarter(journalPublishDate) AS quarter
FROM Practicum2DB.Journal"

tempData8 = dbGetQuery(mydb, sqlCmdData8)



```

```{sql connection=mydb}
SELECT * FROM starschema.PublishDateDimension
       LIMIT 5;
```

### ArticleDateDimension

```{sql connection=mydb}

DROP TABLE IF EXISTS starschema.ArticleDateDimension;

```


```{r}
# construct a SQL query for create
sqlCmdData9 = "CREATE TABLE starschema.ArticleDateDimension (
  articleDatedim_key int NOT NULL AUTO_INCREMENT PRIMARY KEY,
  articleDate DATE,
  day_of_week INT,
year_num INT,
quarter_of_year INT
);"

tempData9 = dbGetQuery(mydb, sqlCmdData9)

# construct a SQL query for insert
sqlCmdData10 = "INSERT INTO starschema.ArticleDateDimension (articleDate,day_of_week,year_num,quarter_of_year)
SELECT DISTINCT articleDate,dayofweek(articleDate) AS day_of_week,year(articleDate) AS year,quarter(articleDate) AS quarter
FROM Practicum2DB.Article"

tempData10 = dbGetQuery(mydb, sqlCmdData10)



```

```{sql connection=mydb}
SELECT * FROM starschema.ArticleDateDimension
       LIMIT 5;
```

### Fact Tables

#### ArticleFact

```{sql connection=mydb}

DROP TABLE IF EXISTS starschema.ArticleFact;

```


```{r}
# construct a SQL query for create
sqlCmdDataf1 = "CREATE TABLE starschema.ArticleFact (
InformationArticleDimensionKey INT,
JournalDimensionKey VARCHAR(20),
pubDatedim_key INT,
articleDatedim_key INT,
mapAuthorDimensionKey INT,
historyDimKey INT,
CONSTRAINT PRIMARY KEY (InformationArticleDimensionKey,JournalDimensionKey,pubDatedim_key,articleDatedim_key,mapAuthorDimensionKey,historyDimKey),
FOREIGN KEY (InformationArticleDimensionKey) REFERENCES starschema.InformationArticleDimension(InformationArticleDimensionKey),
FOREIGN KEY (JournalDimensionKey) REFERENCES starschema.JournalDimension(JournalDimensionKey),
FOREIGN KEY (pubDatedim_key) REFERENCES starschema.PublishDateDimension(pubDatedim_key),
FOREIGN KEY (articleDatedim_key) REFERENCES starschema.ArticleDateDimension(articleDatedim_key),
FOREIGN KEY (mapAuthorDimensionKey) REFERENCES starschema.AuthorMapDimension(mapAuthorDimensionKey),
FOREIGN KEY (historyDimKey) REFERENCES starschema.HistoryDimension(historyDimKey)
)"

tempDataf1 = dbGetQuery(mydb, sqlCmdDataf1)

# construct a SQL query for insert
sqlCmdDatf2 = "INSERT INTO starschema.ArticleFact 
SELECT a.articleId,a.issn,dtdPub.pubDatedim_key, dtdArt.articleDatedim_key, dtdMap.mapAuthorDimensionKey, dtdHis.historyId
   FROM Practicum2DB.Article a
   JOIN Practicum2DB.Journal jd ON jd.issn = a.issn
   JOIN starschema.PublishDateDimension dtdPub ON dtdPub.journalPublishDate = jd.journalPublishDate
   JOIN starschema.ArticleDateDimension dtdArt ON dtdArt.articleDate = a.articleDate
   JOIN starschema.AuthorMapDimension dtdMap ON dtdMap.InformationArticleDimensionKey = a.articleId
   JOIN Practicum2DB.PubMedHistory dtdHis ON dtdHis.articleId = a.articleId"

tempDataf2 = dbGetQuery(mydb, sqlCmdDatf2)



```

```{sql connection=mydb}
SELECT * FROM starschema.ArticleFact
       LIMIT 5;
```

### SUMMARY TABLE



### QUARTER DIMENSION

```{sql connection=mydb}
DROP TABLE IF EXISTS starschema.QuarterDimension;
```


```{r}
# construct a SQL query for create
sqlCmdDataquatcreate = "CREATE TABLE starschema.QuarterDimension (
  quarterYear_key int NOT NULL AUTO_INCREMENT PRIMARY KEY,
  quarter_of_year INT,
  year_num INT
)"

tempDataquatcreate = dbGetQuery(mydb, sqlCmdDataquatcreate)

# construct a SQL query for insert
sqlCmdDataquartIns = "INSERT INTO starschema.QuarterDimension (quarter_of_year,year_num)
SELECT DISTINCT quarter(journalPublishDate) AS quarter,year(journalPublishDate) AS year from Practicum2DB.Journal"

tempDataquartIns = dbGetQuery(mydb, sqlCmdDataquartIns)


```

```{sql connection=mydb}
SELECT * FROM starschema.QuarterDimension
       LIMIT 5;
```

###  SUMMARY TABLE


```{sql connection=mydb}
DROP TABLE IF EXISTS starschema.SummaryFact;
```


```{r}
# construct a SQL query for create
sqlCmdDataSUMMARYCr = "CREATE TABLE starschema.SummaryFact (
summary_key int NOT NULL AUTO_INCREMENT PRIMARY KEY,
totalArticleCount INT,
quarterYear_key INT,
authorDimensionKey INT,
JournalDimensionKey VARCHAR(20),
FOREIGN KEY (JournalDimensionKey) REFERENCES starschema.JournalDimension(JournalDimensionKey),
FOREIGN KEY (quarterYear_key) REFERENCES starschema.QuarterDimension(quarterYear_key),
FOREIGN KEY (authorDimensionKey) REFERENCES starschema.AuthorDimension(authorDimensionKey)
)"

tempDataSummarycreate = dbGetQuery(mydb, sqlCmdDataSUMMARYCr)

# construct a SQL query for insert
sqlCmdDataSummaryIns = "INSERT INTO starschema.SummaryFact (totalArticleCount, quarterYear_key, authorDimensionKey, JournalDimensionKey)
SELECT count(*),dtdQtr.quarterYear_key,dtdMap.authorDimensionKey,a.issn
   FROM Practicum2DB.Article a
   JOIN starschema.JournalDimension jd ON jd.JournalDimensionKey = a.issn
   JOIN starschema.PublishDateDimension dtdPub ON dtdPub.journalPublishDate = jd.journalPublishDate
   JOIN starschema.AuthorMapDimension dtdMap ON dtdMap.InformationArticleDimensionKey = a.articleId
   JOIN starschema.QuarterDimension dtdQtr ON dtdQtr.quarter_of_year = dtdPub.quarter_of_year AND dtdQtr.year_num = dtdPub.year_num
   group by dtdMap.authorDimensionKey,dtdPub.quarter_of_year,a.issn"

tempDataSummaryIns = dbGetQuery(mydb, sqlCmdDataSummaryIns)


```

```{sql connection=mydb}
SELECT * FROM starschema.SummaryFact
       LIMIT 5;
```

## Seasonal graph for the created fact tables

```{r}
library(ggplot2)
```


### Number of journals published per quarter.

```{r}
## Number of journals published per quarter.

#### From the Summary fact table, we use a join with starschema.QuarterDimension to get the quarter of year. This quarter is used to get the distinct journal published per quarter.

tempRsJournalPerQuarter <- dbGetQuery(mydb, statement= paste("select quarter_of_year, count(distinct JournalDimensionKey) as `TotalJournalCount` from starschema.SummaryFact
JOIN  starschema.QuarterDimension using(quarterYear_key)
group by quarter_of_year order by quarter_of_year"))

tempRsJournalPerQuarter

```


```{r}
ggplot(tempRsJournalPerQuarter, aes(x=quarter_of_year, y=TotalJournalCount)) +
  geom_line(color="#69b3a2", size=1) +
  ggtitle("Number of journals published per quarter") +
geom_text(aes(label=TotalJournalCount), vjust=-.5)
```

### Number of authors who featured in a published journal per quarter.

```{r warning=FALSE}
## Number of authors who featured in a published journal per quarter.

#### From the Summary fact table, we use a join with starschema.QuarterDimension to get the quarter of year. This quarter is used to get the authors who featured in the journal published per quarter.


tempRsAuthorsPerQuarter <- dbGetQuery(mydb, statement= paste("select quarter_of_year, SUM(Authcnt) as `TotalAuthor` from(
select  quarter_of_year, count(authorDimensionKey) as Authcnt  from starschema.SummaryFact
JOIN  starschema.QuarterDimension using(quarterYear_key)
group by quarter_of_year,totalArticleCount) as newTable group by quarter_of_year"))

tempRsAuthorsPerQuarter

```



```{r}
ggplot(tempRsAuthorsPerQuarter, aes(x=quarter_of_year, y=TotalAuthor)) +
  geom_line(color="#69b3a2", size=1) +
  ggtitle("Number of authors who featured in a published journal per quarter") +
geom_text(aes(label=TotalAuthor), vjust=-.5)
```

### Number of articles published  per quarter

```{r}

## Number of articles published  per quarter

tempRsArticlesPerQuarter <- dbGetQuery(mydb, statement= paste("select quarter_of_year, count(distinct InformationArticleDimensionKey) as `Total Articles`
from starschema.ArticleFact
JOIN  starschema.PublishDateDimension using(pubDatedim_key) 
group by quarter_of_year"))

tempRsArticlesPerQuarter

```

```{r}
ggplot(tempRsArticlesPerQuarter, aes(x=quarter_of_year, y=`Total Articles`)) +
  geom_line(color="#69b3a2", size=1) +
  ggtitle("Number of articles published per quarter") +
geom_text(aes(label=`Total Articles`), vjust=-.5)
```


### Disconnecting the DB


```{r}
dbDisconnect(mydb)
```